Signatures and Induction Principles for Higher Inductive-Inductive Types

### 2019-02-04 paper.tex
stuff

\item\emph{Iterated equalities and path induction in signatures.}
We can support this relatively easily by closing the universe under equality
type formation, and also using a standard (although size-restricted) definition
of path induction.
\end{itemize}

@@

(5) There is a different equality type former, which can be used to express
paths between type constructors. To our knowledge, this has not been considered
so far in HIT literature, although Cartmell considered equations between sorts
in generalized algebraic theories \cite{gat}. For an example, paths between type
constructors allow a compact definition of integers:
\begin{alignat*}{5}
  & \boldsymbol{\cdot},\,\,Int : U,\,\,zero : \underline{Int},\,p : Int =_\U Int
\end{alignat*}
Here, successor and predecessor functions can be recovered from the $p : Int
=_\U Int$ equality, by transporting an $Int$ along $p$ or $p^{-1}$. This
signature could be unfolded to a larger one, by replacing $p : Int =_\U Int$
with an explicit $suc : Int \ra \underline{Int}$ constructor and additional
constructors expressing that $suc$ is an equivalence. The definition of $Int$
with a successor equivalence was previously suggested by Altenkirch and Pinyo
\cite{pinyo2018integers} and Cavallo and M\"ortberg \cite{cavallointegers}. Our
signature can be viewed as a shorthand for the unfolded version.

@@

(6) We have a function space where the domain is a type in the external
theory. We distinguish it from (3) by using red brick color in the domain
specification. We specify lists and integers as follows, with integers now
given as quotients of pairs of natural numbers:
\begin{alignat*}{4}
  & \hat{A}\in\Type_{\hat{0}} && \vdash\,\boldsymbol{\cdot},\,\, && List:\U,\,\,\,nil:\underline{List},\,\,\,cons:(\hat{x}\in \hat{A})\ra List\ra\underline{List} \\
  & \hat{\Gamma} && \vdash\,\boldsymbol{\cdot},\,\,&& Int:\U,\,\,\,pair:(\hat{x}\,\hat{y}\in\hat{Nat})\ra\underline{Int},\,\,\, \\
  & && && eq:(\hat{a}\,\hat{b}\,\hat{c}\,\hat{d}\in \hat{Nat})(\hat{p}\in \hat{a}\hat{\mathbin{+}}\hat{d}\eqh_{\hat{Nat}} \hat{b}\hat{\mathbin{+}}\hat{c})\ra \underline{pair\,\hat{a}\,\hat{b}=_{Int}\mathsf{pair}\,\hat{c}\,\hat{d}}, \\
  & && && trunc:(x y : Int)(p\,q : a=_{Int} b)\ra \underline{p=_{x=_{Int} y} q}
\end{alignat*}
In the case of integers, $\hat{\Gamma}$ is
$\hat{Nat}\in\Type_{\hat{0}}\hat{,}\,\blank\hat{\mathbin{+}}\blank\in\hat{Nat}\rah
\hat{Nat}\rah\hat{Nat}$, or alternatively, we could require natural
numbers in the external theory. As another example, propositional
truncation for a type $\hat{A}$ is specified as follows.
\[
\hat{A}\in\Type_{\hat{0}}\vdash\, \boldsymbol{\cdot},\,\,\,tr:\U,\,\,\,emb : (\hat{x}\in \hat{A})\ra \underline{tr},\,\,\,eq:(x\,y:tr)\ra \underline{x=_{tr} y}
\]
The smallness of $\hat{A}$ is required in (5). It is possible to generalize
signatures to arbitrary universe levels, but it is not essential to the current
development. Note that the (5) function space preserves strict positivity, since
in the external theory there is no way to recursively refer to the inductive type
\emph{being defined}. The situation is analogous to the case of $W$-types
\cite{abbot05containers}, where shapes and positions can contain arbitrary types
but they cannot recursively refer to the $W$-type being defined.

(7) $\U$ is also closed under a function space where the domain is an external
type and the codomain is a small source theory type. We overload the application
notation for external parameters, as it is usually clear from context which
application is meant. The rules allow types with infinitary constructors, for
example trees containing elements of $\hat{A}$ at the leaves and branching by
$\hat{B}$ (which could be an infinite type):
\[
\hat{A}\in\Type_{\hat{0}}\commah\hat{B}\in\Type_{\hat{0}}\vdash\,\boldsymbol{\cdot},\,\,\,T:\U,\,\,\,leaf:(\hat{x}\in \hat{A})\ra\underline{T},\,\,\,node:((\hat{x}\in \hat{B})\ra T)\ra\underline{T}
\]
Here, $leaf$ has a function type (5) and $node$ has a function type (3) with a
function type (6) in the domain. More generally, we can define $W$-types
\cite{abbot05containers} as follows. $\hat{S}$ describes the ``shapes'' of the
constructors and $\hat{P}$ the ``positions'' where recursive arguments can
appear.
\[
\hat{S}\in\Type_{\hat{0}}\commah\hat{P}:\hat{S}\rah\Type_{\hat{0}}\vdash\,\boldsymbol{\cdot},\,\,\,W:\U,\,\,\,sup: (\hat{s} \in \hat{S})\ra((\hat{p} \in \hat{P}\,\hat{s})\ra W)\ra \underline{W}
\]
For a more complex infinitary example, see the definition of Cauchy
reals in \cite[Definition 11.3.2]{HoTTbook}. It can be also found as
an example file in our Haskell implementation.

@@

\begingroup
\allowdisplaybreaks
\begin{alignat*}{5}
  & \boldsymbol{\cdot}^\S\,\hat{\gamma}\,\hat{\gamma^D} && :\equiv \hat{\top} \\
  & (\Delta,\,x:A)^\S\,(\hat{\gamma},\hat{t})\,(\hat{\gamma^D},\hat{t^D}) && :\equiv \hat{(}\hat{\gamma^S}\in\Delta^\S\hat{\gamma^2}\hat{)}\timesh A^\S\,\hat{\gamma^S}\,\hat{t}\,\hat{t^D} \\
  & x^\S\,\hat{\gamma^S} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma^S} \\
  & \U^\S\,\hat{\gamma^S}\,\hat{A}\,\hat{A^D} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  \hat{A^D}\,\hat{x} \\
  & (\underline{a})^\S\,\hat{\gamma^S}\,\hat{t}\,\hat{t^D} && :\equiv a^\S\,\hat{\gamma^S}\,\hat{t} \eqh \hat{t^D} \\
  & ((x:a)\ra B)^\S\,\hat{\gamma^S}\,\hat{f}\,\hat{f^D} && :\equiv \hat{(}\hat{x}\in a^\A\,\hat{\gamma}\hat{)}\rah  B^\S\,\hat{(}\hat{\gamma,x}\hat{)}\,\hat{(}\hat{\gamma^D,}\,a^\S\,\hat{\gamma^S}\,\hat{x}\hat{)}\,\hat{(}\hat{\gamma^S,}\,\hat{\refl}\hat{)} \\
  & && \hspace{8.4em} \hat{(}\hat{f}\,\hat{x}\hat{)}\,\hat{\big(}\hat{f^D}\,\hat{x}\,\hat{(}a^\S\,\hat{\gamma^S}\,\hat{x}\hat{)}\hat{\big)} \\
  & (t\,u)^\S\,\hat{\gamma^S} && :\equiv \hat{\J}\,\hat{(}t^\S\,\hat{\gamma^S}\,\hat{(} u^\A\,\hat{\gamma}\hat{)}\hat{)}\,\hat{(}u^\S\,\hat{\gamma^S}\hat{)} \\
  & ((\hat{x}\in \hat{A})\ra B)^\S\,\hat{\gamma^S}\,\hat{f}\,\hat{f^D} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah B^\S\,\hat{\gamma^S}\,\hat{(}\hat{f}\,\hat{x}\hat{)}\,\hat{(}\hat{f^D}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^\S\,\hat{\gamma^S} && :\equiv t^\S\,\hat{\gamma^S}\,\hat{u} \\
  & (t=_a u)^\S\,\hat{\gamma^S}\,\hat{e} && :\equiv \TR\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}\hat{\big(}\TR\,\hat{(}u^\S\,\hat{\gamma^S}\hat{)}\,\hat{(}\apd\,\hat{(}a^\S\,\hat{\gamma^S}\hat{)}\,\hat{e}\hat{)}\hat{\big)} \\
  & (\refl_t)^\S\,\hat{\gamma^S} && :\equiv \hat{\J}\,\hat{\refl}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)} \\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^\S\,\hat{\gamma^S} && :\equiv \\
  & && \hspace{-10em}\hat{\J}\,\hat{\bigg(}\hat{\J}\,\hat{\Big(}\hat{\J}\,\hat{\big(}\hat{\J}\,\hat{(}\hat{\lambda}\,\hat{p^D}\,\hat{p^S}\,\hat{pr^D}\,\hat{pr^S}\hat{.}\,\hat{pr^S}\hat{)}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}\,\\
  & && \hspace{-9em}\hat{(}\hat{\mathsf{uncurry}}\,p^\D\,\hat{\gamma^2}\hat{)}\,\hat{(}\hat{\mathsf{uncurry}}\,p^\S\,\hat{\gamma^S}\hat{)}\,\hat{(}pr^\D\,\hat{\gamma^2}\hat{)}\,\hat{(}pr^\S\,\hat{\gamma^S}\hat{)}\hat{\big)}\,\hat{(}eq^\A\,\hat{\gamma}\hat{)}\hat{\Big)}\,\hat{(}u^\S\,\hat{\gamma^S}\hat{)}\hat{\bigg)}\,\hat{(}eq^\S\,\hat{\gamma^S}\hat{)} \\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^\S\,\hat{\gamma^S} && :\equiv \\
  & && \hspace{-10em}\hat{\J}\,\hat{\big(}\hat{\J}\,\hat{(}\hat{\lambda}\,\hat{p^D}\,\hat{p^S}\hat{.}\,\hat{\refl}\hat{)}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}\,\hat{(}\hat{\mathsf{uncurry}}\,p^\D\,\hat{\gamma^2}\hat{)}\,\hat{(}\hat{\mathsf{uncurry}}\,p^\S\,\hat{\gamma^S}\hat{)}\hat{\big)}\,\hat{(}pr^\S\,\hat{\gamma^S}\hat{)}\\
  & ((\hat{x}\in \hat{A})\ra b)^\S\,\hat{\gamma^S\,f\,t} && :\equiv b^\S\,\hat{\gamma^S}\,\hat{(}\hat{f}\,\hat{t}\hat{)} \\
  & (t\,\hat{u})^\S\,\hat{\gamma^S} && :\equiv \ap\,\hat{(}\hat{\lambda} \hat{f.}\hat{f}\,\hat{u}\hat{)}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}
\end{alignat*}
\endgroup

@@

For $a =_\U b$ equalities, let us consider here only the $(a =_\U b)^\S$ case
for the type former. The interpretation here is inspried by the univalence
axiom, although the translation does not require it. If we assume univalence,
then $\hat{a =_{\Type_i} b}$ equalities in the external theory are equivalent to
equivalences, which contain a function $\hat{f : a \ra b}$ and a proof
$\hat{\mathsf{isEqv}\,f}$. Since $\hat{\mathsf{isEqv}}$ is propositional (proof
irrelevant), we only need to say something about the underlying $\hat{f}$
function in the translation.

Let us consider $\blank^\S$ for higher inductive integers with the type
equality:
\begin{alignat*}{5}
  & && (\boldsymbol{\cdot},\,\,Int : U,\,\,zero : \underline{Int},\,p : Int =_\U Int)^\S\,
     \hat{(\tt,\,Int,\,z,\,p)\,(\tt,\,Int^D,\,z^D,\,p^D)} \\
  & \equiv \,\, && \hat{
     \top \times (Int^S : (x : Int)\ra Int^D\,x) \times (z^S : Int^S\,z = z^D)}\\
  & && \hat{\hspace{1em}\times\,
       (p^S : (\lambda\,x.\,p^s\,(\coe\,p\,x)) = (\lambda\,x.\,\TR\,p^D\,(\J\,(p^s\,x)\,p)))}
\end{alignat*}

In Appendix \ref{sec:app}, we show how the type of displayed algebra sections is
computed for the two-dimensional sphere. In Appendix \ref{sec:wtypes}, we show
the same for indexed W-types.

>>>>

\item\emph{Iterated equalities and type constructor equalities with path
indcution.} We can support iterated equalities of point and path constructors
by closing the universe under equality type formation of point and path
constructors, together with a standard (although size-restricted) definition
of path induction. We can also introduce a different type former for
equalities between type constructors.

\end{itemize}

@@

(5) There is a different equality type former, which can be used to express
paths between type constructors. To our knowledge, this has not been considered
in previous HIT literature, although Cartmell considered equations between sorts
in generalized algebraic theories \cite{gat}. For an example, paths between type
constructors allow a compact definition of integers:
\begin{alignat*}{5}
  & \boldsymbol{\cdot},\,\,Int : U,\,\,zero : \underline{Int},\,p : Int =_\U Int
\end{alignat*}
Here, successor and predecessor functions can be recovered from the $p : Int
=_\U Int$ equality, by transporting an $Int$ along $p$ or $p^{-1}$. This
signature could be unfolded to a larger one, by replacing $p : Int =_\U Int$
with an explicit $suc : Int \ra \underline{Int}$ constructor and additional
constructors expressing that $suc$ is an equivalence. The definition of $Int$
with a successor equivalence was previously suggested by Altenkirch and Pinyo
\cite{pinyo2018integers} and Cavallo and M\"ortberg \cite{cavallointegers}. Our
signature can be viewed as a shorthand for the unfolded version.

@@

(6) We have a function space where the domain is a type in the external
theory. We distinguish it from (3) by using red brick color in the domain
specification. We specify lists and integers as follows, with integers now
given as quotients of pairs of natural numbers:
\begin{alignat*}{4}
  & \hat{A}\in\Type_{\hat{0}} && \vdash\,\boldsymbol{\cdot},\,\, && List:\U,\,\,\,nil:\underline{List},\,\,\,cons:(\hat{x}\in \hat{A})\ra List\ra\underline{List} \\
  & \hat{\Gamma} && \vdash\,\boldsymbol{\cdot},\,\,&& Int:\U,\,\,\,pair:(\hat{x}\,\hat{y}\in\hat{Nat})\ra\underline{Int},\,\,\, \\
  & && && eq:(\hat{a}\,\hat{b}\,\hat{c}\,\hat{d}\in \hat{Nat})(\hat{p}\in \hat{a}\hat{\mathbin{+}}\hat{d}\eqh_{\hat{Nat}} \hat{b}\hat{\mathbin{+}}\hat{c})\ra \underline{pair\,\hat{a}\,\hat{b}=_{Int}\mathsf{pair}\,\hat{c}\,\hat{d}}, \\
  & && && trunc:(x y : Int)(p\,q : a=_{Int} b)\ra \underline{p=_{x=_{Int} y} q}
\end{alignat*}
In the case of integers, $\hat{\Gamma}$ is
$\hat{Nat}\in\Type_{\hat{0}}\hat{,}\,\blank\hat{\mathbin{+}}\blank\in\hat{Nat}\rah
\hat{Nat}\rah\hat{Nat}$, or alternatively, we could require natural
numbers in the external theory. As another example, propositional
truncation for a type $\hat{A}$ is specified as follows.
\[
\hat{A}\in\Type_{\hat{0}}\vdash\, \boldsymbol{\cdot},\,\,\,tr:\U,\,\,\,emb : (\hat{x}\in \hat{A})\ra \underline{tr},\,\,\,eq:(x\,y:tr)\ra \underline{x=_{tr} y}
\]
The smallness of $\hat{A}$ is required in (6). It is possible to generalize
signatures to arbitrary universe levels, but it is not essential to the current
development. Note that the (6) function space preserves strict positivity, since
in the external theory there is no way to recursively refer to the inductive type
\emph{being defined}. The situation is analogous to the case of $W$-types
\cite{abbot05containers}, where shapes and positions can contain arbitrary types
but they cannot recursively refer to the $W$-type being defined.

(7) $\U$ is also closed under a function space where the domain is an external
type and the codomain is a small source theory type. We overload the application
notation for external parameters, as it is usually clear from context which
application is meant. The rules allow types with infinitary constructors, for
example trees containing elements of $\hat{A}$ at the leaves and branching by
$\hat{B}$ (which could be an infinite type):
\[
\hat{A}\in\Type_{\hat{0}}\commah\hat{B}\in\Type_{\hat{0}}\vdash\,\boldsymbol{\cdot},\,\,\,T:\U,\,\,\,leaf:(\hat{x}\in \hat{A})\ra\underline{T},\,\,\,node:((\hat{x}\in \hat{B})\ra T)\ra\underline{T}
\]
Here, $leaf$ has a function type (6) and $node$ has a function type (3) with a
function type (7) in the domain. More generally, we can define $W$-types
\cite{abbot05containers} as follows. $\hat{S}$ describes the ``shapes'' of the
constructors and $\hat{P}$ the ``positions'' where recursive arguments can
appear.
\[
\hat{S}\in\Type_{\hat{0}}\commah\hat{P}:\hat{S}\rah\Type_{\hat{0}}\vdash\,\boldsymbol{\cdot},\,\,\,W:\U,\,\,\,sup: (\hat{s} \in \hat{S})\ra((\hat{p} \in \hat{P}\,\hat{s})\ra W)\ra \underline{W}
\]
For a more complex infinitary example, see the definition of Cauchy
reals in \cite[Definition 11.3.2]{HoTTbook}. It can be also found as
an example file in our Haskell implementation.

@@

\begingroup
\allowdisplaybreaks
\begin{alignat*}{5}
  & \boldsymbol{\cdot}^\S\,\hat{\gamma}\,\hat{\gamma^D} && :\equiv \hat{\top} \\
  & (\Delta,\,x:A)^\S\,(\hat{\gamma},\hat{t})\,(\hat{\gamma^D},\hat{t^D}) && :\equiv \hat{(}\hat{\gamma^S}\in\Delta^\S\hat{\gamma^2}\hat{)}\timesh A^\S\,\hat{\gamma^S}\,\hat{t}\,\hat{t^D} \\
  & x^\S\,\hat{\gamma^S} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma^S} \\
  & \U^\S\,\hat{\gamma^S}\,\hat{A}\,\hat{A^D} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  \hat{A^D}\,\hat{x} \\
  & (\underline{a})^\S\,\hat{\gamma^S}\,\hat{t}\,\hat{t^D} && :\equiv a^\S\,\hat{\gamma^S}\,\hat{t} \eqh \hat{t^D} \\
  & ((x:a)\ra B)^\S\,\hat{\gamma^S}\,\hat{f}\,\hat{f^D} && :\equiv \hat{(}\hat{x}\in a^\A\,\hat{\gamma}\hat{)}\rah  B^\S\,\hat{(}\hat{\gamma,x}\hat{)}\,\hat{(}\hat{\gamma^D,}\,a^\S\,\hat{\gamma^S}\,\hat{x}\hat{)}\,\hat{(}\hat{\gamma^S,}\,\hat{\refl}\hat{)} \\
  & && \hspace{8.4em} \hat{(}\hat{f}\,\hat{x}\hat{)}\,\hat{\big(}\hat{f^D}\,\hat{x}\,\hat{(}a^\S\,\hat{\gamma^S}\,\hat{x}\hat{)}\hat{\big)} \\
  & (t\,u)^\S\,\hat{\gamma^S} && :\equiv \hat{\J}\,\hat{(}t^\S\,\hat{\gamma^S}\,\hat{(} u^\A\,\hat{\gamma}\hat{)}\hat{)}\,\hat{(}u^\S\,\hat{\gamma^S}\hat{)} \\
  & (t=_a u)^\S\,\hat{\gamma^S}\,\hat{e} && :\equiv \TR\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}\hat{\big(}\TR\,\hat{(}u^\S\,\hat{\gamma^S}\hat{)}\,\hat{(}\apd\,\hat{(}a^\S\,\hat{\gamma^S}\hat{)}\,\hat{e}\hat{)}\hat{\big)} \\
  & (\refl_t)^\S\,\hat{\gamma^S} && :\equiv \hat{\J}\,\hat{\refl}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)} \\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^\S\,\hat{\gamma^S} && :\equiv \\
  & && \hspace{-10em}\hat{\J}\,\hat{\bigg(}\hat{\J}\,\hat{\Big(}\hat{\J}\,\hat{\big(}\hat{\J}\,\hat{(}\hat{\lambda}\,\hat{p^D}\,\hat{p^S}\,\hat{pr^D}\,\hat{pr^S}\hat{.}\,\hat{pr^S}\hat{)}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}\,\\
  & && \hspace{-9em}\hat{(}\hat{\mathsf{uncurry}}\,p^\D\,\hat{\gamma^2}\hat{)}\,\hat{(}\hat{\mathsf{uncurry}}\,p^\S\,\hat{\gamma^S}\hat{)}\,\hat{(}pr^\D\,\hat{\gamma^2}\hat{)}\,\hat{(}pr^\S\,\hat{\gamma^S}\hat{)}\hat{\big)}\,\hat{(}eq^\A\,\hat{\gamma}\hat{)}\hat{\Big)}\,\hat{(}u^\S\,\hat{\gamma^S}\hat{)}\hat{\bigg)}\,\hat{(}eq^\S\,\hat{\gamma^S}\hat{)} \\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^\S\,\hat{\gamma^S} && :\equiv \\
  & && \hspace{-10em}\hat{\J}\,\hat{\big(}\hat{\J}\,\hat{(}\hat{\lambda}\,\hat{p^D}\,\hat{p^S}\hat{.}\,\hat{\refl}\hat{)}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}\,\hat{(}\hat{\mathsf{uncurry}}\,p^\D\,\hat{\gamma^2}\hat{)}\,\hat{(}\hat{\mathsf{uncurry}}\,p^\S\,\hat{\gamma^S}\hat{)}\hat{\big)}\,\hat{(}pr^\S\,\hat{\gamma^S}\hat{)}\\
  & (a=_\U b)^\S\,\hat{\gamma^S} && :\equiv
    \hat{\lambda\,e\,e^D.\,(\lambda\,x.\, \blc{b^\S}\,\gamma^S\, (\coe\,e\,x))
     =
    (\lambda\,x.\, \TR\,e^D\,(\J\,(\blc{a^\S}\,\gamma^S\,x)\,e))}\\
  & (\refl_a)^\S\,\hat{\gamma^S} && :\equiv \hat{\refl} \\
  & (\J_{a\,(x.z.p)}\,pr\,_b\,eq)^\S\,\hat{\gamma^S} && :\equiv\\
  & && \hspace{-10em}\hat{\J\,\bigg(\lambda\,b^D\,\,b^S\,eq^D.\,\J\,\Big(\lambda\,b^S\,eq^S.\,\J\,(\lambda P^S\,pr^S.\,pr^S)\,eq^S\,(\mathsf{uncurry}\,\blc{p^\S}\,\gamma^S)\,(\blc{pr^\S}\,\gamma^S)\Big)}  \\
  & && \hspace{-9em}\hat{\,(\blc{eq^\D}\,\gamma^D)\,b^S\bigg)\,(\blc{eq^\A}\,\gamma)\,(\blc{b^\D}\,\gamma^D)\,(\blc{b^\S}\,\gamma^S)\,(\blc{eq^\D}\,\gamma^D)\,(\blc{eq^\S}\,\gamma^S) } \\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^\S\,\hat{\gamma^S} && :\equiv \hat{\J\,\refl\,(\blc{pr^\S}\,\gamma^S)} \\
  & ((\hat{x}\in \hat{A})\ra B)^\S\,\hat{\gamma^S}\,\hat{f}\,\hat{f^D} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah B^\S\,\hat{\gamma^S}\,\hat{(}\hat{f}\,\hat{x}\hat{)}\,\hat{(}\hat{f^D}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^\S\,\hat{\gamma^S} && :\equiv t^\S\,\hat{\gamma^S}\,\hat{u} \\
  & ((\hat{x}\in \hat{A})\ra b)^\S\,\hat{\gamma^S\,f\,t} && :\equiv b^\S\,\hat{\gamma^S}\,\hat{(}\hat{f}\,\hat{t}\hat{)} \\
  & (t\,\hat{u})^\S\,\hat{\gamma^S} && :\equiv \ap\,\hat{(}\hat{\lambda} \hat{f.}\hat{f}\,\hat{u}\hat{)}\,\hat{(}t^\S\,\hat{\gamma^S}\hat{)}
\end{alignat*}
\endgroup

@@

For $a =_\U b$ equalities, let us examine here only the $(a =_\U b)^\S$ case.
The interpretation here is inspired by the univalence axiom, although the
translation does not require it. If we assume univalence in the external theory,
then $\hat{a =_{\Type_i} b}$ equalities in are equivalent to equivalences, which
contain $\hat{f : a \ra b}$ functions together with $\hat{\mathsf{isEqv}\,f}$
proofs. Hence, we can view the $\hat{e : \blc{a^\A}\,\gamma =
  \blc{a^\A}\,\gamma}$ and $\hat{e^D : \TR_{(\lambda\,A.\,A \ra
    \mathsf{Type_0})}\,e\,(\blc{a^\D}\,\gamma^D) = \blc{b^\D}\,\gamma^D}$ proofs
as function bundled together with witnesses of equivalence. Since
$\hat{\mathsf{isEqv}}$ is proof irrelevant, we only need to relate the
underlying functions. For $\hat{e}$, the underlying function is given just by
$\hat{\coe\,e}$. For $\hat{e^D}$, the underlying function is a bit more
complicated, since $\hat{e^D}$ is a path over a path. We can observe this in the
example for higher inductive integers:
\begin{alignat*}{5}
  & && (\boldsymbol{\cdot},\,\,Int : U,\,\,zero : \underline{Int},\,p : Int =_\U Int)^\S\,
     \hat{(\tt,\,Int,\,z,\,p)\,(\tt,\,Int^D,\,z^D,\,p^D)} \\
  & \equiv \,\, && \hat{
     \top \times (Int^S : (x : Int)\ra Int^D\,x) \times (z^S : Int^S\,z = z^D)}\\
  & && \hat{\hspace{1em}\times\,
       (p^S : (\lambda\,x.\,p^S\,(\coe\,p\,x)) = (\lambda\,x.\,\TR\,p^D\,(\J\,(p^S\,x)\,p)))}
\end{alignat*}

In Appendix \ref{sec:app}, we additionally show how the type of displayed
algebra sections is computed for the two-dimensional sphere. In Appendix
\ref{sec:wtypes}, we show the same for indexed W-types.
